How Kids Code and How We Know:\\An Exploratory Study on the Scratch Repository

### 2016-04-22 Paper/paper.tex
Changes based on Sohon's comments
Smaller figure margins
Added comment on temporary githib link

\maketitle
\begin{abstract}
Block-based programming languages like Scratch, Alice and Blockly are becoming increasingly common as introductory languages in programming education. There is substantial research showing that these visual programming environments are suitable for teaching programming concepts. But, what do people do when they use Scratch? In this paper we explore the characteristics of Scratch programs. To this end we have scraped the Scratch public repository and retrieved 250.000 projects. We present an analysis of these projects in three different dimensions. Initially, we look at the types of blocks used and the size of the projects. We then investigate complexity, used abstractions and programming concepts. Finally we detect \emph{code smells} like large scripts, dead code and duplicated code blocks. Our results show that 1) most Scratch programs are small, however Scratch programs consisting of over 100 sprites exist, 2) programming abstraction concepts like procedures are not commonly used and 3) Scratch programs do suffer from code smells including large scripts and unmatched broadcast signals.
\end{abstract}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{fig/ui.png}
		\caption{The Scratch user interface consisting of the `cat' sprite on the left, the toolbox with available blocks in the category `Events' in the middle and the code associated with the sprite on the right.}
		\label{fig:ui}
	\end{center}
\end{figure}

@@

We ran Kragle on March 2nd 2016 for 24 hours and, in that time, it obtained a little over 250,000 projects. Out of the \nPrograms, we failed to parse and further analyze 2,367 projects due to technical difficulties with the provided JSON files. Kragle, as well as all scraped projects and our analysis files are available.\footnote{\label{repo}\url{https://github.com/ScratchLover42/ICER-Data-Code}}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/11cloneslines}
	\caption{Number of cloned scripts of different block sizes across and within sprites}
	\label{fig:cloneslines}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/13longmethod}
	\caption{Size of sprites and scripts in number of blocks}
	\label{fig:longmethod}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1sprites}
	\caption{Number of sprites in the analyzed projects}
	\label{fig:size-sprites}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1scripts}
	\caption{Number of scripts in the analyzed projects}
	\label{fig:size-scripts}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1locs}
	\caption{Number of blocks in the analyzed projects}
	\label{fig:size-locs}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/3cyclomatic}
	\caption{McCabe cyclomatic complexity of the \nscripts~analyzed scripts}
	\label{fig:cyclomatic}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/16Categories}
	\caption{Number of blocks from each category in the analyzed projects}
	\label{fig:categories}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/16Shapes}
	\caption{Number of blocks of each shape in the analyzed projects}
	\label{fig:shapes}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/6proceduresperproject}
	\caption{Number of procedures for the 17,979 projects that include at least one}
	\label{fig:proceduresperproject}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/5arguments}
	\caption{Number of arguments for the procedures in the dataset}
	\label{fig:procedurearguments}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/7procedurecalls}
	\caption{Number of calls of each procedure in the dataset}
	\label{fig:procedurecalls}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/8variableslists}
	\caption{Number of variables and lists used in the projects}
	\label{fig:variables}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/11clonesprojects}
	\caption{Cloned scripts in the dataset projects}
	\label{fig:clonesprojects}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/11clonescopies}
	\caption{Number of copies of the identified clones}
	\label{fig:clonescopies}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/12deadcode}
	\caption{Size of scripts identified as dead code}
	\label{fig:deadcode}
\end{figure}

>>>>

\maketitle
\begin{abstract}
Block-based programming languages like Scratch, Alice and Blockly are becoming increasingly common as introductory languages in programming education. There is substantial research showing that these visual programming environments are suitable for teaching programming concepts. But, what do people do when they use Scratch? In this paper we explore the characteristics of Scratch programs. To this end we have scraped the Scratch public repository and retrieved 250,000 projects. We present an analysis of these projects in three different dimensions. Initially, we look at the types of blocks used and the size of the projects. We then investigate complexity, used abstractions and programming concepts. Finally we detect \emph{code smells} like large scripts, dead code and duplicated code blocks. Our results show that 1) most Scratch programs are small, however Scratch programs consisting of over 100 sprites exist, 2) programming abstraction concepts like procedures are not commonly used and 3) Scratch programs do suffer from code smells including large scripts and unmatched broadcast signals.
\end{abstract}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{fig/ui.png}
		\vspace{-1em}
		\caption{The Scratch user interface consisting of the `cat' sprite on the left, the toolbox with available blocks in the category `Events' in the middle and the code associated with the sprite on the right.}
		\label{fig:ui}
	\end{center}
\end{figure}

@@

We ran Kragle on March 2nd 2016 for 24 hours and, in that time, it obtained a little over 250,000 projects. Out of the \nPrograms, we failed to parse and further analyze 2,367 projects due to technical difficulties with the provided JSON files. Kragle, as well as all scraped projects and our analysis files are available.\footnote{\label{repo}\url{https://github.com/ScratchLover42/ICER-Data-Code} (temporary link to protect the anonymity of the authors)}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/11cloneslines}
	\vspace{-1em}
	\caption{Number of cloned scripts of different block sizes across and within sprites}
	\label{fig:cloneslines}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/13longmethod}
	\vspace{-1em}
	\caption{Size of sprites and scripts in number of blocks}
	\label{fig:longmethod}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1sprites}
	\vspace{-1em}
	\caption{Number of sprites in the analyzed projects}
	\label{fig:size-sprites}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1scripts}
	\vspace{-1em}
	\caption{Number of scripts in the analyzed projects}
	\label{fig:size-scripts}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1locs}
	\vspace{-1em}
	\caption{Number of blocks in the analyzed projects}
	\label{fig:size-locs}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/3cyclomatic}
	\vspace{-1em}
	\caption{McCabe cyclomatic complexity of the \nscripts~analyzed scripts}
	\label{fig:cyclomatic}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/16Categories}
	\vspace{-1em}
	\caption{Number of blocks from each category in the analyzed projects}
	\label{fig:categories}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/16Shapes}
	\vspace{-1em}
	\caption{Number of blocks of each shape in the analyzed projects}
	\label{fig:shapes}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/6proceduresperproject}
	\vspace{-1em}
	\caption{Number of procedures for the 17,979 projects that include at least one}
	\label{fig:proceduresperproject}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/5arguments}
	\vspace{-1em}
	\caption{Number of arguments for the procedures in the dataset}
	\label{fig:procedurearguments}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/7procedurecalls}
	\vspace{-1em}
	\caption{Number of calls of each procedure in the dataset}
	\label{fig:procedurecalls}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/8variableslists}
	\vspace{-1em}
	\caption{Number of variables and lists used in the projects}
	\label{fig:variables}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/11clonesprojects}
	\vspace{-1em}
	\caption{Cloned scripts in the dataset projects}
	\label{fig:clonesprojects}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/11clonescopies}
	\vspace{-1em}
	\caption{Number of copies of the identified clones}
	\label{fig:clonescopies}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/12deadcode}
	\vspace{-1em}
	\caption{Size of scripts identified as dead code}
	\label{fig:deadcode}
\end{figure}
